Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2022-08-12 Supplemental-Material-A-Annotated-Analysis.Rmd
Update Supplemental-Material-A-Annotated-Analysis.Rmd

#################################################################
##               EXPORT FOR STM MACHINE LEARNING               ##
#################################################################
# bannerCommenter::banner("EXPORT FOR STM MACHINE LEARNING", centre = TRUE, bandChar = "#")
varNamPosAdd <- c("relatedness", "autonomy", "competence")

dtAllStm <- rbind(
  dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>%
    mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
  dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
  dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V")
  ) %>%
  select(ID,
         PID,
         study,
         TID,
         TIDnum,
         date,
         week,
         IntergroupContact,
         IngroupContact,
         everything(),
         -starts_with("ResponseId")) %>%
  as.data.frame

dtAllStm2 <- dtAllStm

dtAllStm2$IntGrAnx.pre_calc <-
  psych::scoreItems(
    keys = c(rep(1,6),rep(-1,6)),
    items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
    min = 1,
    max = 10
  )$scores %>%
  as.numeric

dtAllStm2$IntGrAnx.post_calc <-
  psych::scoreItems(
    keys = c(rep(1,6),rep(-1,6)),
    items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
    min = 1,
    max = 10
  )$scores %>%
  as.numeric

dtAllStm2$swl.pre_calc <-
  psych::scoreItems(
    keys = rep(1,5),
    items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
    min = 1,
    max = 7,
    missing = TRUE,
    impute = "none",
  )$scores %>%
  as.numeric

dtAllStm2$discrimination.post_calc <-
  psych::scoreItems(
    keys = rep(1,9),
    items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
    min = 1,
    max = 6,
    missing = TRUE,
    impute = "none",
  )$scores %>%
  as.numeric

dtAllStm3 <- dtAllStm2 %>%
  select(
    -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
    -matches("^IntGrAnx[0-9]{2}.?\\.post"),
    -matches("^SWL[0-9]{2}\\.pre"),
    -matches("^EvDayDiscr[0-9]{2}\\.post"),
    -relatednessSelf,
    -relatednessOther,
    -relatednessNoInteraction
  )

write.csv2(dtAllStm3,
           file = "data/keyMotivesSTM.csv")
#################################################################

@@

@@

@@

t3MainFull <- ThreeWay::T3(
  dtMainImpWide, 
  paste("PP", 1:nrow(dtMainImpWide), sep = "_"), 
  names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
  paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
)

t3Main <- ThreeWay::T3func(
  X = dtMainImpWide,
  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
  m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
  p = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
  r1 = 3,
  r2 = 3,
  r3 = 2,
  start = 0,
  conv = 1e-6
)

FitT3Main <- T3fitpartitioning(
  Xprep = dtMainImpWide,
  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
  m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
  p = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
  AS = t3Main$A, 
  BT = t3Main$B, 
  CU = t3Main$C, 
  K = t3Main$H, 
  renormmode = 0,
  laba = paste("PP", 1:nrow(dtMainImpWide), sep = "_"), 
  labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
  labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
)

capture.output(
  threeWayAnoBiDailyImp <- threewayanova(
    Y = dtMainImpWide,
    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
    m = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
    p = length(unique(dtPcaS123Imp$TIDnum)) # Nr. time points (C)
  ), 
file='NUL')

rrcov3way::Tucker3(
  X = dtMainImpWide, 
  P = nrow(dtMainImpWide), # Nr. ppt.s (A)
  Q = length(names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars]), # Nr. variables (B)
  R = length(unique(dtPcaS123Imp$TIDnum)), # Nr. time points (C)
  center = FALSE, #center.mode = c("A", "B", "C", "AB", "AC", "BC", "ABC"), 
  scale = FALSE, #scale.mode = c("B", "A", "C"), 
  conv = 1e-06, start="svd",
  robust = FALSE, coda.transform=c("none", "ilr", "clr"), 
  ncomp.rpca = 0, alpha = 0.75, robiter=100, crit=0.975, trace = FALSE
)

>>>>

#################################################################
##               EXPORT FOR STM MACHINE LEARNING               ##
#################################################################
# bannerCommenter::banner("EXPORT FOR STM MACHINE LEARNING", centre = TRUE, bandChar = "#")
# varNamPosAdd <- c("relatedness", "autonomy", "competence")
# 
# dtAllStm <- rbind(
#   dtS1$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S1") %>%
#     mutate(across(!TID & !study & !KeyNeed & !DaytimeNeed, as.numeric)),
#   dtS2$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S2"),
#   dtS3$full %>% select(KeyNeed, DaytimeNeed, any_of(varNamS123MiRed), any_of(varNamPosAdd)) %>% mutate(study = "S3")
# ) %>%
#   group_by(study, PID) %>%
#   mutate(ID = cur_group_id()) %>%
#   ungroup %>%
#   mutate(
#     date = as.Date(gsub(" .*", "", TID)),
#     week = strftime(date, format = "%Y-W%V")
#   ) %>%
#   select(ID,
#          PID,
#          study,
#          TID,
#          TIDnum,
#          date,
#          week,
#          IntergroupContact,
#          IngroupContact,
#          everything(),
#          -starts_with("ResponseId")) %>%
#   as.data.frame
# 
# dtAllStm2 <- dtAllStm
# 
# dtAllStm2$IntGrAnx.pre_calc <-
#   psych::scoreItems(
#     keys = c(rep(1,6),rep(-1,6)),
#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.pre")),
#     min = 1,
#     max = 10
#   )$scores %>%
#   as.numeric
# 
# dtAllStm2$IntGrAnx.post_calc <-
#   psych::scoreItems(
#     keys = c(rep(1,6),rep(-1,6)),
#     items = dtAllStm2 %>% select(matches("^IntGrAnx[0-9]{2}.?\\.post")),
#     min = 1,
#     max = 10
#   )$scores %>%
#   as.numeric
# 
# dtAllStm2$swl.pre_calc <-
#   psych::scoreItems(
#     keys = rep(1,5),
#     items = dtAllStm2 %>% select(matches("^SWL[0-9]{2}\\.pre")),
#     min = 1,
#     max = 7,
#     missing = TRUE,
#     impute = "none",
#   )$scores %>%
#   as.numeric
# 
# dtAllStm2$discrimination.post_calc <-
#   psych::scoreItems(
#     keys = rep(1,9),
#     items = dtAllStm2 %>% select(matches("^EvDayDiscr[0-9]{2}\\.post")),
#     min = 1,
#     max = 6,
#     missing = TRUE,
#     impute = "none",
#   )$scores %>%
#   as.numeric
# 
# dtAllStm3 <- dtAllStm2 %>%
#   select(
#     -matches("^IntGrAnx[0-9]{2}.?\\.pre"),
#     -matches("^IntGrAnx[0-9]{2}.?\\.post"),
#     -matches("^SWL[0-9]{2}\\.pre"),
#     -matches("^EvDayDiscr[0-9]{2}\\.post"),
#     -relatednessSelf,
#     -relatednessOther,
#     -relatednessNoInteraction
#   )
# 
# write.csv2(dtAllStm3,
#            file = "data/keyMotivesSTM.csv")
#################################################################

@@

1.  Center across participants (within time point and variable).\
2.  Normalize across time points and participants (within variable).\
3.  Select number of components for each mode (i.e., mode complexity; using Convex Hull procedure heuristic based on total number of components and scree test value taken from a PCA-SUP analysis [where, PCASUP analyses are PCA's of supermatrices with slices of the 3way array next to each other, thus 3 supermatrices are analyed by PCA, and for each we get a component matrix, and eigenvalues. The eigenvalues may give an indication as to the required number of components for each mode.]).
4.  Run 3MPCA
5.  Test

@@

capture.output(
  threeWayAnoBiDailyImp <- threewayanova(
    Y = dtMainImpWide,
    n = nrow(dtMainImpWide), # Nr. ppt.s (A)
    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
    p = length(unique(sub("_.*", "", names(dtMainImpWide)))) # Nr. time points (C)
  ), 
file='NUL')

@@

# t3MainFull <- ThreeWay::T3(
#   dtMainImpWide,
#   paste("PP", 1:nrow(dtMainImpWide), sep = "_"),
#   names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars],
#   paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
# )

dtMainImpWideC <- ThreeWay::cent3(
  X = dtMainImpWide, 
  n = nrow(dtMainImpWide), # Nr. ppt.s (A)
  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWide)))), # Nr. variables (B)
  p = length(unique(sub("_.*", "", names(dtMainImpWide)))), # Nr. time points (C)
  mode = 1
) %>%
  as.data.frame
#apply(dtMainImpWideC,2,sum)

dtMainImpWideCZ <- ThreeWay::norm3(
  X = dtMainImpWideC, 
  n = nrow(dtMainImpWideC), # Nr. ppt.s (A)
  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideC)))), # Nr. variables (B)
  p = length(unique(sub("_.*", "", names(dtMainImpWideC)))), # Nr. time points (C)
  mode = 2
) %>%
  as.data.frame
names(dtMainImpWideCZ) <- names(dtMainImpWideC)
#apply(dtMainImpWideC,2,sum)
# dtMainImpWideCZ %>%
#   select(ends_with("_AttitudesDutch")) %>%
#   unlist %>%
#   sd
# psych::describe(dtMainImpWideCZ)

MainPcaSup <- ThreeWay::T3runsApproxFit(
  X = dtMainImpWideCZ, 
  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
  maxa = 20,
  maxb = 20,
  maxc = 20
)
MainDimSelector <- capture.output(
  ThreeWay::DimSelector(
    out = MainPcaSup,
    n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
    m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
    p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
    model = 2
  )
)

MainDim <- tail(MainDimSelector, n = 1) %>%
  str_extract("\\(.*\\)") %>%
  stringr::str_extract_all('\\d+',simplify=TRUE) %>%
  as.data.frame %>% 
  mutate_all(as.numeric) %>%
  dplyr::rename(c(
    A = V1,
    B = V2,
    C = V3
  ))
MainDim

t3Main <- ThreeWay::T3func(
  X = dtMainImpWideCZ,
  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
  r1 = MainDim$A,
  r2 = MainDim$B,
  r3 = MainDim$C,
  start = 0,
  conv = 1e-6
)

FitT3Main <- T3fitpartitioning(
  Xprep = dtMainImpWideCZ,
  n = nrow(dtMainImpWideCZ), # Nr. ppt.s (A)
  m = length(unique(sub("T[0-9]{2}_", "", names(dtMainImpWideCZ)))), # Nr. variables (B)
  p = length(unique(sub("_.*", "", names(dtMainImpWideCZ)))), # Nr. time points (C)
  AS = t3Main$A, 
  BT = t3Main$B, 
  CU = t3Main$C, 
  K = t3Main$H, 
  renormmode = 0,
  laba = paste("PP", 1:nrow(dtMainImpWideCZ), sep = "_"), 
  labb = names(dtPcaS123Imp)[!names(dtPcaS123Imp) %in% idVars], 
  labc = paste("T", 1:length(unique(dtPcaS123Imp$TIDnum)), sep = "")
)

t3MainRot <- varimcoco(
  A = t3Main$A, 
  B = t3Main$B, 
  C = t3Main$C, 
  H = t3Main$H,
  wa_rel = 0,
  wb_rel = 0,
  wc_rel = 0
)

plot(t3MainRot$C[, 1], type = "n", ylim = range(t3MainRot$C), xlab = "Time occasion", ylab = "Component score")
points(t3MainRot$C[, 1], pch = 16)
lines(t3MainRot$C[, 1])
points(t3MainRot$C[, 2], pch = 17)
lines(t3MainRot$C[, 2])
legend("topleft", legend = c("C1", "C2"), pch = c(16, 17), bty = "n")

# library(factoextra)
# fviz_nbclust(t3MainRot$A[,1:2], kmeans, method = "wss")
# clusterdata <- t3MainRot$A[,1:2]
# km.res <- kmeans(clusterdata, 6, nstart = 25)
# fviz_cluster(object = km.res, # kmeans object 
#              data = clusterdata, # data used for clustering
#              ellipse.type = "norm",
#              geom = "point",
#              palette = "jco",
#              main = "",
#              xlab = "PC1",
#              ylab = "PC2",
#              ggtheme = theme_minimal())

# lares::corr_cross(
#   dtMainImpWideCZ, # name of dataset
#   max_pvalue = 0.05, # display only significant correlations (at 5% level)
#   top = 50 # display top 50 couples of variables (by correlation coefficient)
# )
